Wigglyhedra

### 2024-02-26 draft/wigglyhedra.tex
minor corrections, Cambrian fan, associahedron

% geometry
\newcommandx{\Fan}[1][1=D]{\mathcal{F}_{#1}} % fan
\newcommand{\polytope}[1]{\mathds{#1}} % font polytope
\newcommandx{\Zono}[1][1=D]{\polytope{Z}_{#1}} % zonotope
\newcommandx{\Asso}[1][1=n]{\polytope{A}_{#1}} % associahedron

% specific wiggly
\newcommand{\wigglyComplex}{\mathrm{WC}} % wiggly complex
\newcommand{\wigglyFlipGraph}{\mathrm{WFG}} % wiggly flip graph
\newcommand{\wigglyIncreasingFlipGraph}{\mathrm{WIFG}} % wiggly increasing flip graph
\newcommand{\wigglyLattice}{\mathrm{WL}} % wiggly lattice
\newcommand{\wigglyFan}{\mathrm{WF}} % wiggly fan
\newcommand{\wigglyhedron}{\polytope{W}} % wigglyhedron

@@

If~$\alpha$ and~$\alpha'$ are non pointed with~$i' = j$, then we have
\[
\alpha^+ \sqcup \alpha'^+ = \beta^+ \sqcup \{2j-1, 2j\} = \beta'^+
\qquad\text{and}\qquad 
\alpha^- \sqcup \alpha'^- = \beta^- = \beta'^- \sqcup \{2j-1, 2j\}.
\]
We thus obtain that
\[
\one_{\alpha^+} + \one_{\alpha'^+} = \one_{\beta^+} + \b{e}_{2j-1} + \b{e}_{2j} = \one_{\beta'^+}
\qquad\text{and}\qquad 
\one_{\alpha^-} + \one_{\alpha'^-} = \one_{\beta^-} = \one_{\beta'^-} + \b{e}_{2j-1} + \b{e}_{2j}.
\]
hence
\begin{align*}
2 \big( \hat{\b{g}}(\alpha) + \hat{\b{g}}(\alpha') \big)
& = 2(\one_{\alpha^+} - \one_{\alpha^-}) + 2(\one_{\alpha'^+} - \one_{\alpha'^-})
= 2(\one_{\alpha^+} + \one_{\alpha'^+}) - 2(\one_{\alpha^-} + \one_{\alpha'^-}) \\
& = (\one_{\beta^+} + \one_{\beta'^+} + \b{e}_{2j-1} + \b{e}_{2j}) - (\one_{\beta^-} + \one_{\beta'^-} + \b{e}_{2j-1} + \b{e}_{2j}) \\
& = (\one_{\beta^+} + \one_{\beta'^+}) - (\one_{\beta^-} + \one_{\beta'^-})
= (\one_{\beta^+} - \one_{\beta^-}) + (\one_{\beta^-} + \one_{\beta'^-})
= \hat{\b{g}}(\beta) + \hat{\b{g}}(\beta').
\end{align*}

If~$\alpha$ and~$\alpha'$ are crossing and there is~$k \in [n-1]$ with $k \in (A \cap B') \cup (\{i\} \cap B') \cup (A \cap \{i'\})$ and~$k+1 \in (A' \cap B) \cup (\{j'\} \cap B) \cup (A' \cap \{j\})$, then we have
\begin{gather*}
\alpha^+ \cup \alpha'^+ = \beta^+ \cup \beta'^+ \qquad\text{and}\qquad \alpha^+ \cap \alpha'^+ = \beta^+ \cap \beta'^+, \\
\alpha^- \cup \alpha'^- = \beta^- \cup \beta'^- \qquad\text{and}\qquad \alpha^- \cap \alpha'^- = \beta^- \cap \beta'^-.
\end{gather*}
We thus obtain that
\[
\one_{\alpha^+} + \one_{\alpha'^+} = \one_{\alpha^+ \cup \alpha'^+} + \one_{\alpha^+ \cap \alpha'^+} = \one_{\beta^+ \cup \beta'^+} + \one_{\beta^+ \cap \beta'^+} = \one_{\beta^+} + \one_{\beta'^+}
\]
and similarly~$\one_{\alpha^-} + \one_{\alpha'^-} = \one_{\beta^-} + \one_{\beta'^-}$, so that~$\hat{\b{g}}(\alpha) + \hat{\b{g}}(\alpha') =\hat{\b{g}}(\beta) + \hat{\b{g}}(\beta')$.
\end{proof}

@@

\begin{definition}
The \defn{$\delta$-gon} is a convex $(n+2)$-gon with a point at abscissa~$i$ and ordinate of the sign of~$\delta_i$  for all~$0 \le i \le n+1$
A \defn{$\delta$-triangulation} is a set of ordered pairs~$(i,j)$ which are the internal diagonals of a triangulation of the $\delta$-gon.
The \defn{$\delta$-triangulation lattice} is the transitive closure of the slope increasing flip graph on $\delta$-triangulations.
%If~$T$ and~$T'$ are two $\delta$-triangulations with~$T \ssm \{d\} = T' \ssm \{d'\}$, then the flip from~$T$ to~$T'$ is \defn{increasing} if the slope of~$d$ is smaller than the slope of~$d'$.
\end{definition}

@@

\begin{definition}
\label{def:CambrianWigglyPseudotriangulations}
The \defn{$\delta$-wiggly arcs} are the arcs~$(0, j, {[1,j[}, \varnothing)$ for~$\delta_j > 0$ and~$(0, j, \varnothing, {[1,j[})$ for~$\delta_j < 0$.
A \defn{$\delta$-wiggly pseudotriangulation} is a wiggly pseudotriangulation containing all $\delta$-wiggly arcs.
The \defn{$\delta$-wiggly pseudotriangulation lattice} is the transitive closure of the increasing flip graph on $\delta$-wiggly pseudotriangulations.
\end{definition}

@@

\begin{proposition}
\label{prop:CambrianIntervals}
For any signature~$\delta \in \{+,-\}^n$, the $\delta$-triangulation lattice, $\delta$-permutation lattice, \mbox{$\delta$-wiggly} pseudotriangulation lattice, and $\delta$-wiggly permutation lattice are all isomorphic, and are known as the \defn{$\delta$-Cambrian lattice}~\cite{Reading-CambrianLattices}.
\end{proposition}

@@

\begin{remark}
We had to make a choice in \cref{def:CambrianWigglyPseudotriangulations,def:CambrianWigglyPermutations}.
For~$\delta_j = {+}$, we could replace ``$(0, j, {[1,j[}, \varnothing)$'' by~``$(j, n+1, {]j,n]}, \varnothing)$'' in \cref{def:CambrianWigglyPseudotriangulations}, and ``$\sigma^{-1}(i) \le \sigma^{-1}(2j-1)$ for~$1 \le i \le 2j$'' by ``$\sigma^{-1}(i) \le \sigma^{-1}(2j-1)$ for~$2j-1 \le i \le 2n$'' in \cref{def:CambrianWigglyPermutations}.
Independently, for~$\delta_j = {-}$, we could replace ``$(0, j, \varnothing, {[1,j[})$'' by~``$(j, n+1, \varnothing, {]j,n]})$'' in \cref{def:CambrianWigglyPseudotriangulations}, and ``$\sigma^{-1}(2j) \le \sigma^{-1}(i)$ for~$1 \le i \le 2j$'' by ``$\sigma^{-1}(2j-1) \le \sigma^{-1}(i)$ for~$2j-1 \le i \le 2n$'' in \cref{def:CambrianWigglyPermutations}.
%\end{itemize}
\end{remark}

\vincent{todo: find an explicit affine map between the Hohlweg-Lange $\delta$-associahedron and the face of the wigglyhedron corresponding to all $\delta$-wiggly arcs.}

>>>>

% geometry
\newcommandx{\Fan}[1][1=D]{\mathcal{F}_{#1}} % fan
\newcommand{\polytope}[1]{\mathds{#1}} % font polytope

% specific wiggly
\newcommand{\wigglyComplex}{\mathrm{WC}} % wiggly complex
\newcommand{\wigglyFlipGraph}{\mathrm{WFG}} % wiggly flip graph
\newcommand{\wigglyIncreasingFlipGraph}{\mathrm{WIFG}} % wiggly increasing flip graph
\newcommand{\wigglyLattice}{\mathrm{WL}} % wiggly lattice
\newcommand{\wigglyFan}{\mathrm{WF}} % wiggly fan
\newcommand{\wigglyhedron}{\polytope{W}} % wigglyhedron
\newcommand{\Asso}{\polytope{A}\mathsf{sso}} % associahedron

@@

If~$\alpha$ and~$\alpha'$ are non pointed, then we have
\[
\alpha^+ \sqcup \alpha'^+ = \beta^+ \sqcup \{2j-1, 2j\} = \beta'^+
\qquad\text{and}\qquad 
\alpha^- \sqcup \alpha'^- = \beta^- = \beta'^- \sqcup \{2j-1, 2j\}.
\]
We thus obtain that
\[
\one_{\alpha^+} + \one_{\alpha'^+} = \one_{\beta^+} + \b{e}_{2j-1} + \b{e}_{2j} = \one_{\beta'^+}
\qquad\text{and}\qquad 
\one_{\alpha^-} + \one_{\alpha'^-} = \one_{\beta^-} = \one_{\beta'^-} + \b{e}_{2j-1} + \b{e}_{2j}.
\]
hence
\begin{align*}
2 \big( \hat{\b{g}}(\alpha) + \hat{\b{g}}(\alpha') \big)
& = 2(\one_{\alpha^+} - \one_{\alpha^-}) + 2(\one_{\alpha'^+} - \one_{\alpha'^-})
= 2(\one_{\alpha^+} + \one_{\alpha'^+}) - 2(\one_{\alpha^-} + \one_{\alpha'^-}) \\
& = (\one_{\beta^+} + \one_{\beta'^+} + \b{e}_{2j-1} + \b{e}_{2j}) - (\one_{\beta^-} + \one_{\beta'^-} + \b{e}_{2j-1} + \b{e}_{2j}) \\
& = (\one_{\beta^+} + \one_{\beta'^+}) - (\one_{\beta^-} + \one_{\beta'^-})
= (\one_{\beta^+} - \one_{\beta^-}) + (\one_{\beta^-} + \one_{\beta'^-})
= \hat{\b{g}}(\beta) + \hat{\b{g}}(\beta').
\end{align*}

If~$\alpha$ and~$\alpha'$ cross, then we have
\begin{gather*}
\alpha^+ \cup \alpha'^+ = \beta^+ \cup \beta'^+ \qquad\text{and}\qquad \alpha^+ \cap \alpha'^+ = \beta^+ \cap \beta'^+, \\
\alpha^- \cup \alpha'^- = \beta^- \cup \beta'^- \qquad\text{and}\qquad \alpha^- \cap \alpha'^- = \beta^- \cap \beta'^-.
\end{gather*}
We thus obtain that
\[
\one_{\alpha^+} + \one_{\alpha'^+} = \one_{\alpha^+ \cup \alpha'^+} + \one_{\alpha^+ \cap \alpha'^+} = \one_{\beta^+ \cup \beta'^+} + \one_{\beta^+ \cap \beta'^+} = \one_{\beta^+} + \one_{\beta'^+}
\]
and similarly~$\one_{\alpha^-} + \one_{\alpha'^-} = \one_{\beta^-} + \one_{\beta'^-}$, so that~$\hat{\b{g}}(\alpha) + \hat{\b{g}}(\alpha') =\hat{\b{g}}(\beta) + \hat{\b{g}}(\beta')$.
\end{proof}

@@

\begin{definition}
The \defn{$\delta$-gon} is a convex $(n+2)$-gon with a point with abscissa~$i$ and with ordinate of the same sign as~$\delta_i$  for all~$0 \le i \le n+1$.
A \defn{$\delta$-triangulation} is a triangulation of the $\delta$-gon.
The \defn{$\delta$-triangulation lattice} is the transitive closure of the slope increasing flip graph on $\delta$-triangulations.
%If~$T$ and~$T'$ are two $\delta$-triangulations with~$T \ssm \{d\} = T' \ssm \{d'\}$, then the flip from~$T$ to~$T'$ is \defn{increasing} if the slope of~$d$ is smaller than the slope of~$d'$.
\end{definition}

@@

\begin{definition}
\label{def:CambrianWigglyPseudotriangulations}
The \defn{$\delta$-wiggly arcs} are the arcs~$(0, j, {[1,j[}, \varnothing)$ for~$\delta_j = {+}$ and~$(0, j, \varnothing, {[1,j[})$ for~$\delta_j = {-}$.
A \defn{$\delta$-wiggly pseudotriangulation} is a wiggly pseudotriangulation containing all $\delta$-wiggly arcs.
The \defn{$\delta$-wiggly pseudotriangulation lattice} is the transitive closure of the increasing flip graph on $\delta$-wiggly pseudotriangulations.
\end{definition}

@@

\begin{remark}
We had to make a choice in \cref{def:CambrianWigglyPseudotriangulations,def:CambrianWigglyPermutations}.
For~$\delta_j = {+}$, we could replace ``$(0, j, {[1,j[}, \varnothing)$'' by~``$(j, n+1, {]j,n]}, \varnothing)$'' in \cref{def:CambrianWigglyPseudotriangulations}, and ``$\sigma^{-1}(i) \le \sigma^{-1}(2j-1)$ for~$1 \le i \le 2j$'' by ``$\sigma^{-1}(i) \le \sigma^{-1}(2j-1)$ for~$2j-1 \le i \le 2n$'' in \cref{def:CambrianWigglyPermutations}.
Independently, for~$\delta_j = {-}$, we could replace ``$(0, j, \varnothing, {[1,j[})$'' by~``$(j, n+1, \varnothing, {]j,n]})$'' in \cref{def:CambrianWigglyPseudotriangulations}, and ``$\sigma^{-1}(2j) \le \sigma^{-1}(i)$ for~$1 \le i \le 2j$'' by ``$\sigma^{-1}(2j-1) \le \sigma^{-1}(i)$ for~$2j-1 \le i \le 2n$'' in \cref{def:CambrianWigglyPermutations}.
\cref{prop:CambrianLattice,prop:CambrianFan} would still hold.
\end{remark}

\begin{proposition}
\label{prop:CambrianLattice}
For any signature~$\delta \in \{+,-\}^n$, the $\delta$-triangulation lattice, $\delta$-permutation lattice, \mbox{$\delta$-wiggly} pseudotriangulation lattice, and $\delta$-wiggly permutation lattice are all isomorphic, and are known as the \defn{$\delta$-Cambrian lattice}~\cite{Reading-CambrianLattices}.
\end{proposition}

@@

Recall that the $\delta$-Cambrian lattice can be realized by the $\delta$-Cambrian fan~\cite{ReadingSpeyer} and the $\delta$-associahedron of~\cite{HohlwegLange}.
Namely, to each diagonal~$(i,j)$ of the $\delta$-gon, we associate the subset~$X(\delta)$ of points of the $\delta$-gon distinct from~$0$ and~$n+1$, and located below the line~$(i,j)$, including~$i$ if and only if~$\delta_i = {+}$ and~$j$ if and only if~$\delta_j = {+}$.
The \defn{$\delta$-Cambrian fan} is the fan with a ray~$\b{g}(d) \eqdef \pi(\one_{X(d)})$ for each internal diagonal~$d$ of the $\delta$-gon and a cone~$\R_{\ge0} \setangle{\b{g}(d)}{d \in D^\circ}$ for each dissection~$D$ of the $\delta$-gon.
The \defn{$\delta$-associahedron} is the polytope defined by the equality~$\dotprod{\one_n}{\b{x}} = \binom{n+1}{2}$ and the inequalities~$\dotprod{\b{g}(d)}{\b{x}} \le \binom{|X(d)|+1}{2}$ for all internal diagonals~$d$ of the $\delta$-gon.

\begin{proposition}
\label{prop:CambrianFan}
The $\delta$-associahedron~$\Asso_\delta$ is normally equivalent to the face of the wigglyhedron~$\wigglyhedron_n$ corresponding to the wiggly pseudodissection formed by the $\delta$-wiggly arcs.
\end{proposition}

\begin{proof}
The maps of \cref{prop:CambrianLattice} preserve the $\b{g}$-vector linear dependences.
\vincent{explain}
\end{proof}

\vincent{todo: find an explicit affine map between the Hohlweg-Lange $\delta$-associahedron and the face of the wigglyhedron corresponding to all $\delta$-wiggly arcs.}
\vincent{todo: add picture}
